### 2016-12-18 preamble.tex
Copied russian text

\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian, english]{babel}

\usepackage{comment}

\usepackage{enumerate}
\usepackage{cite}
\usepackage{mathrsfs}
\usepackage{fullpage}
\usepackage{hyperref} % \url{url} \href{url}{displayed text}
\usepackage{tabularx}
\usepackage{indentfirst}
\usepackage{bm}
\usepackage{braket}

\usepackage{csquotes}
\MakeOuterQuote{"}
\usepackage[pdftex]{graphicx}
\graphicspath{{pic/}}
\usepackage{tikz}
\usepackage{xcolor} % \textcolor
\usepackage{mathtools} % xmapsto, \Vector
\usepackage{comment}
\usepackage{soul} % text striking
\usepackage{amsmath,amsfonts,amssymb,amsthm,amscd,amsxtra}
\usepackage{physics}
\usepackage[sharp]{easylist}
\usepackage[colorinlistoftodos,prependcaption,disable]{todonotes} % TODO: move to preamble!!
\presetkeys%
    {todonotes}%
    {inline,backgroundcolor=red}{}
\newcommand{\mtodo}[1]{\todo[linecolor=green!70!white, backgroundcolor=blue!20!white, bordercolor=red]{#1}}