### 2016-12-18 paper.tex

\input{preamble.tex}
\input{macros/macros.tex}

\begin{document}

\input{bundle.tex}
\end{document}