### 2016-12-20 paper.tex
Inlined macros, less packages

\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian, english]{babel}

% \usepackage{fullpage}
\usepackage{cite}
\usepackage{hyperref} % \url{url} \href{url}{displayed text}

\usepackage[pdftex]{graphicx}
\graphicspath{{pic/}}
\usepackage{tikz}
\usepackage{xcolor} % \textcolor
\usepackage{mathtools} % xmapsto, \Vector
\usepackage{amsmath,amsfonts,amssymb,amsthm,amscd,amsxtra}
\usepackage{physics}
\usepackage[colorinlistoftodos,prependcaption,disable]{todonotes} % TODO: move to preamble!!
\presetkeys%
    {todonotes}%
    {inline,backgroundcolor=red}{}
\newcommand{\mtodo}[1]{\todo[linecolor=green!70!white, backgroundcolor=blue!20!white, bordercolor=red]{#1}}

\newcommand{\hilb}[1]{\mathcal{H}_{#1}}
\newcommand{\cconj}[1]{\overline{#1}}
\newcommand{\hank}[1]{H_{#1}^{(1)}}

\newcommand{\mcF}{\mathcal{F}}
\newcommand{\mcH}{\mathcal{H}} % ???
\newcommand{\mcI}{\mathcal{I}}
\newcommand{\mcL}{\mathcal{L}} % L2 space
\newcommand{\mcO}{\mathcal{O}} % Big O
\newcommand{\mcP}{\mathcal{P}}

\newcommand{\bbC}{\mathbb{C}} % complex plane
\newcommand{\bbD}{\mathbb{D}} % complex unit disk
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbK}{\mathbb{K}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbT}{\mathbb{T}} % complex unit circle
\newcommand{\bbZ}{\mathbb{Z}}

% imaginary unit
\newcommand{\iu}{{i\mkern1mu}}

\newcommand{\eps}{\varepsilon}

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\phi}{\varphi} % TODO is there a prettier way to do that

\newcommand{\eexp}[1]{e^{#1}}

\DeclareMathOperator\atanh{atanh}

\begin{document}

\input{bundle.tex}
\end{document}